A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-01-09 paper.Rmd
more on neonatal

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path='figs/',fig.height=5,comment=NA)
library(knitr)
library(eha)
g <- function(x) (log(1 + x))^3 # x in days.
ginv <- function(y) exp(y^(1/3)) - 1
```

@@

## The Bourgeois-Pichat procedure

@@

## Post-neonatal mortality

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path='figs/',fig.height=4,comment=NA)
library(knitr)
library(eha)
g <- function(x) (log(1 + x))^3 # x in days.
ginv <- function(y) exp(y^(1/3)) - 1
```

@@

### The Bourgeois-Pichat procedure

@@

### Post-neonatal mortality

@@

So, the *hazard function* (which is the slope of the lines in Figure \@ref(fig:pneo))
is simply estimated by the ratio of the number of deaths to the total time of exposure:

```{r simpest, echo = TRUE}
posthaz <- with(ginfdat, sum(event) / sum(exit - enter))
```

which evaluates to `r format(posthaz, digits = 3, scientific = TRUE)`. A small number, but remember
that the time unit is defined on the scale defined by the function $g$ (close to *day*).

Is the sum of two components, *external* and *internal* causes. According to the *B-P*
theory, the external component is equal to the post-neonatal mortality, which contains 
only an external component.

We create the neonatal data set by censor all lives at the age 28:

```{r neon, echo = TRUE}
neo <- age.window(infdat, c(0, 28))
neo$enter <- g(neo$enter)
neo$exit <- g(neo$exit)
```

and then fit a piecewise constant hazards model with cutpoints at (0, 1, 2, 4, 8, 12, 18, 24, 30, 35, 39), see 
Figure \@ref(fig:pchneo).

```{r pchneo, fig.cap = "Neonatal hazard function with external component (dashed).", fig.height = 4}
fitneo <- pchreg(Surv(enter, exit, event) ~ 1, data = neo, cuts = c(0,  1, 2, 4, 8, 12, 18, 24, 30, 35, 39))
plot(fitneo, fn = "haz", col = "blue", main = "", xlab = "g(days)")
abline(h = 0, v = 0)
abline(h = posthaz, col = "red", lty = 2)
```

Let us take a closer look at what happens at the end of the neonatal inerval:

```{r pchneo2, fig.cap = "Neonatal hazard function with external component (dashed), right hand tail.", fig.height = 4}
##fitneo <- pchreg(Surv(enter, exit, event) ~ 1, data = neo, cuts = c(0,  1, 2, 4, 8, 12, 18, 24, 30, 39))
plot(fitneo, fn = "haz", col = "blue", main = "", xlab = "g(days)", xlim = c(15, 39))
abline(h = 0, v = 0)
abline(h = posthaz, col = "red", lty = 2)
```

A smooth transition to the post-neonatal mortality.
